Paper_Results

### 2025-02-13 Paper_Results.Rmd
update files

---
title: "Paper_Examples"
output: html_document
date: "2025-01-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(R.matlab)
library(foreach)
library(doParallel)
library(dplyr)
library(lme4)
library(tseries)
library(caret)
library(pROC)
library(boot)
library(patchwork)
```

@@

# Basic Parameters from Study
n_features <- 10 # FIXED IN SAEB
n_samples <- 90  # Timepoints/ RECORDS (fixed in saeb) Number of samples per subject
n_subjects <- 150 # SUBJECTS (VARIED 4 TO 32) # Needs to be an even number

# Generating Outcome
overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
sqrt(overall_prob_outcome*(1-overall_prob_outcome))
sd_outcome <- 0.2 # Controls variability BETWEEN different subject (e.g., tracking depression 0.21)

@@

source("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Simulation_Functions_notshiny.R")

set.seed(12361488)
features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect)
features_sample <- features_sample[[1]]

@@

result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")

@@

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example2.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

@@

```{r}
############# Figure 5 #################
########################################
result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")

@@

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example4.jpg",
  plot = combined_plot,
  width = 10.2,
  height = 5.5)

mean(result$auc)
mean(result$auc_c)

mean(result$auc_c_individual)
mean(result$auc_individual)

test <- result$auc_c-result$auc
sum((test > 0)/length(test))

unique(result$A)
```

```{r}
############# Figure 6 #################
########################################
result <- read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results.csv")
result <- result[result$A != 0 & result$A != 0.8,]
p1 <- ggplot(result, aes(x=auc_c, y=auc_c_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person centered") +
  xlab("AUC centered") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5)   + 
annotate("text", x = 0.5, y = 0.9, label = "Linear relationship between overall AUC \nand within-person AUC",  hjust = 0) +
  geom_curve(
    aes(x = 0.63, y = 0.87, xend = 0.75, yend = 0.8), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.15
  ) +
  ggtitle("Uncentered Predictors")

@@

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example3.jpg",
  plot = combined_plot,
  width = 11.5,
  height = 6.5)
```

@@

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/Example_var.jpg",
  plot = p2,
  width = 8,
  height = 7)

@@

```{r }
############ Table 5 #############
##################################

sim_1$auc_value_base # row-wise
sim_4$auc_value_base # rolling window

@@

result = read.csv("/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/simulation_results_baseline.csv")
p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
  geom_point(alpha=0.5, size = 1.2, color = "#83AF9B") +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#83AF9B")) +
  ylab("Overall Performance (AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  guides(col = guide_colourbar()) +
  ggtitle(paste("Performance of a Random Intercept Only Model (No Predictors Included)")) +
  geom_hline(yintercept = 0.5) +
  ylim(range= c(0.5,1)) +
    annotate("text", x = 0.5, y = 0.82, label =  "Performance increases as between person differences explain a \nlarger % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.7, y = 0.85, xend = 0.5, yend = 0.91), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  ) +
      annotate("text", x = 0.2, y = 0.62, label =  "Performance can be moderate even if between person differences explain only a relatively \nsmall % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.19, y = 0.62, xend = 0.07, yend = 0.61), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  )

ggsave(  "/Users/f007qrc/Library/CloudStorage/GoogleDrive-anna.m.langener@dartmouth.edu/My Drive/Darmouth Drive/ML CrossValidation Project/figure7.jpg",
  plot = p7,
  width = 10,
  height = 5)

>>>>

---
title: "Paper_Results"
output: html_document
date: "2025-01-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(dplyr)
library(lme4)
library(caret)
library(pROC)
library(patchwork)
```
## General Information
This R Markdown file can be used to replicate our results that are presented in the paper. Please note that it does not contain the shiny app.

@@

# Basic Parameters from Study
n_features <- 10  # Features
n_samples <- 90  # Timepoints
n_subjects <- 150 # Subjects

# Generating Outcome
overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
sd_outcome <- 0.2 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))

@@

source("Simulation_Functions_notshiny.R")

set.seed(12361488)
features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect = FALSE)
features_sample <- features_sample[[1]]

@@

result <- read.csv("simulation_results.csv") # see Run_Simulation.R to investigate how results were created

@@

ggsave("Figures/Example2.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

@@

```{r}
############# Figure 5 #################
########################################
result <- read.csv("simulation_results.csv")

@@

ggsave(  "Figures/Example4.jpg",
  plot = combined_plot,
  width = 10.2,
  height = 5.5)

```{r}
############# Figure 6 #################
########################################
result <- read.csv("simulation_results.csv")
result <- result[result$A != 0 & result$A != 0.8,]
p1 <- ggplot(result, aes(x=auc_c, y=auc_c_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person centered") +
  xlab("AUC centered") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5)   + 
annotate("text", x = 0.5, y = 0.9, label = "Linear relationship between overall AUC \nand within-person AUC",  hjust = 0) +
  geom_curve(
    aes(x = 0.63, y = 0.87, xend = 0.75, yend = 0.8), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.15
  ) +
  ggtitle("Uncentered Predictors")

@@

ggsave( "Figures/Example3.jpg",
  plot = combined_plot,
  width = 11.5,
  height = 6.5)
```

@@

ggsave( "Figures/Example_var.jpg",
  plot = p2,
  width = 8,
  height = 7)

@@

```{r }
############ Table 5 #############
##################################
sim_1$auc_value_base # row-wise
sim_4$auc_value_base # rolling window

@@

result = read.csv("simulation_results_baseline.csv")
p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
  geom_point(alpha=0.5, size = 1.2, color = "#83AF9B") +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#83AF9B")) +
  ylab("Overall Performance (AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  guides(col = guide_colourbar()) +
  ggtitle(paste("Performance of a Random Intercept Only Model (No Predictors Included)")) +
  geom_hline(yintercept = 0.5) +
  ylim(range= c(0.5,1)) +
    annotate("text", x = 0.5, y = 0.82, label =  "Performance increases as between person differences explain a \nlarger % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.7, y = 0.85, xend = 0.5, yend = 0.91), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  ) +
      annotate("text", x = 0.2, y = 0.62, label =  "Performance can be moderate even if between person differences explain only a relatively \nsmall % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.19, y = 0.62, xend = 0.07, yend = 0.61), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  )

ggsave(  "Figures/figure7.jpg",
  plot = p7,
  width = 10,
  height = 5)
